Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions

### 2025-09-24 Files/paper.tex
Previous, with the pdf

\abstract{Identifying the generating mechanism of a real-life network is
difficult as, more often than not, only snapshots are available, but not
the full evolution. One candidate for the generating mechanism is
preferential attachment which, in its simplest form, results in a degree
distribution that follows the power law. Consequently, the growth of
real-life networks that display such power-law behaviour is commonly
modelled by preferential attachment. Recently, the power law has been
challenged by the presence of alternatives with comparable performance,
as well as the recent findings that the tail of the degree distribution
is often lighter than implied by the body, whilst still being regularly
varying. We use methods from discrete extemes to study a version of the
model with a flexible preference function that allows super/sub-linear
behaviour whilst guaranteeing the limiting degree distribution is
regularly varying. We relate the distributions tail index directly to
the model parameters and infer their true values from the degree
distribution alone.}

>>>>

\abstract{Identifying the generating mechanism of a network is
challenging as, more often than not, only snapshots are available, but
not the full evolution. One candidate for the generating mechanism is
preferential attachment which, in its simplest form, results in a degree
distribution that follows the power law. Consequently, the growth of
real-life networks that display such power-law behaviour is commonly
modelled by preferential attachment. The power law has been challenged
by the presence of alternatives with comparable performance, as well as
the recent findings that the tail of the degree distribution is often
lighter than implied by the body, whilst still being regularly varying.
We use methods from discrete extremes to study a version of the model
with a flexible preference function that allows super/sub-linear
behaviour whilst guaranteeing the limiting degree distribution is
regularly varying. We relate the tail index directly to the model
parameters and enable inference of parameters from the degree
distribution alone.}
